Analysis of HTTPS Adoption and Third-party Tracking: \\A Swedish Perspective

### 2015-04-29 paper/paper.tex
Added abstract draft from Niklas

Blah blah

>>>>

Both active third-party tracking (e.g., using scripts and plugins embedded within visited websites) and passive tracking

(e.g., through network monitoring) are extensively used to gather knowledge about users internet activities and interests.

While HTTPS can be used to protect against some privacy concerns raised by passive monitoring, prior work have not carefully

characterized the HTTPS adoption and third-party tracking when using HTTP and HTTPS, respectively.  This paper presents a

measurement analysis of the current HTTPS adoption and third-party tracking usage.  First, we develop and release the

software of a novel measurement framework for automated, repeatable retrieval, and analysis of websites.  Second,

we characterize the HTTPS adoption, HTTPS redirection dynamics, and use of secure/non-secure third-party resources

from both a national (Swedish) and global perspective.  In total, we analyze more than 150,000 websites.  Third,

we characterize the third-party tracking usage and the companies that may get access to user information when using

HTTP and HTTPS.  Overall, we have found that users may be given a false sense of privacy when using HTTPS.

They are often directed to non-secure subdomains, and even when served using HTTPS, the use of third-party

(external) resources is high among all classes of domains, regardless if HTTP or HTTPS is used.  In fact,

third-party usage is most prevalent among the most popular website, visited most often.  While Google has

by far the greatest third-party coverage (more than 90\% for most website categories), we have found that

there are many other domains serving third-party content that may fly under the radar.

%Both active third-party tracking (e.g., using scripts and plugins embedded within visited websites) and passive tracking (e.g., through network monitoring) are extensively used to collect information and knowledge about users internet activities and interests.  While HTTPS can be used to protect against privacy concerns raised by passive monitoring, prior work have not carefully studied the adoption of HTTPS and how much privacy protection against third-party tracking can be gained from using HTTPS.  This paper presents a measurement analysis of the current HTTPS adoption and third-party tracking usage.  First, we develop and release the software of a novel measurement framework for automated, repeatable retrieval, and analysis of websites and their third-party usage and HTTPS dynamics.  Second, we characterize the HTTPS adoption and use of secure/non-secure third-party resources from both a national (Swedish) and global perspective.  In total, we repeatedly analyze more than 150,000 websites.  Third, we characterize the third-party tracking usage and the companies that may get access to user information when using HTTP and HTTPS.  Overall, we have found that users may be given a false sense of privacy when using HTTPS.  They are often directed to non-secure subdomains, and even when served using HTTPS, the use of third-party (external) resources is high among all classes of domains, regardless if HTTP or HTTPS is used.  This is magnified by third-party usage being most prevalent among the most popular website, which people visit most often.  While Google has by far the greatest third-party coverage (more than 90\% for most website categories) among known tracker companies, we have found that there are many other domains that also server third-party content which may fly under the radar.
